---
title: "Statistical Inference Project, Part 1"
author: "Nannette Spear"
date: "July 23, 2018"
output:
  html_document:
    fig_caption: yes
    fig_height: 4
    fig_width: 6
    keep_md: yes
  pdf_document: default
---

##Overview
Part 1 of the course project involves applying the Central Limit
Theorem to exponential distribution in R. The distribution of
averages of 40 exponentials, simulated 1,000 times will be investigated.

- Lambda is 0.2
- Theoretical Mean is 1/0.2
- Theoretical Standard Deviation is also 1/0.2
- Observed data substitutes for the population
- Random sampling (simulations) corresponds to the observations

Load knitr package
```{r load knitr, echo=FALSE}
library(knitr)
```
Set the Working Directory
````{r set working directory, echo=FALSE}
opts_knit$set(root.dir = "C:/Coursera/Statistical Inference/Course Project")
````

Exponential Distribution of 40 exponentials
````{r exponential, echo=TRUE}
x <- rexp(40, 0.20)
n <- length(x)
````

1000 Simulations
````{r simulations, echo=TRUE}
B <- 1000
resamples <- matrix(sample(x, n*B, replace = TRUE),B, n)  
means <- apply(resamples, 1, mean)
````

Theoretical Mean of the distribution
````{r tmean, echo=TRUE}
tmean <- 1/0.20
tmean
````
Estimated Mean of all samples
````{r est mean, echo=TRUE}
estmean <- mean(means)
estmean
````
The theoretical and estimated mean are very close to each other
````{r varmean, echo=TRUE}
meanvar <- estmean - tmean
meanvar
````
The Theroetical Standard Deviation is = 1/lambda/sqrt(n)
````{r tsd, echo = TRUE}
tsd <- (1/0.20)/sqrt(n)
tsd
````
The variance is the standard deviation squared
````{r varsd, echo = TRUE}
tvar <- tsd^2
tvar
````
The estimated standard deviation
````{r estsd, echo = TRUE}
estsd <- sd(means)
estsd
````
The estimated variance is the standard deviation squared
````{r estvar, echo = TRUE}
estvar <- estsd^2
estvar
````
The theoretical and estimated variance are also very close to each other
````{r varvar, echo = TRUE}
varvar <- estvar - tvar
varvar
````

95% Confidence Interval
````{r confidence interval, echo=TRUE}
CI <- quantile(means, c(0.025, 0.975))
CI
````

This plot is an estimate of the sampling distribution of the mean
````{r histogram, echo=TRUE}
library(ggplot2)
g <- ggplot(data.frame(means = means), aes(x=means))
g <- g + geom_histogram(color = "black", fill = "lightblue", binwidth = 0.15)
g <- g + geom_vline(xintercept = 5, size = 1, color = "orange")       
g <- g + geom_vline(xintercept = CI, size = 0.5, color = "black", linetype ="dashed")
g
````

##Conclusion
- The histogram represents the simulation means distributed over
a normal distribution.  
- The orange vertical line at mean 5, represents the 
theoretical mean.  
- The area between the two dashed lines represent where the estimated or sample
means will fall, 95% of the time.

