---
title: "Statistical Inference Project, Part 2"
author: "Nannette Spear"
date: "July 23, 2018"
output: 
  html_document: 
    fig_caption: yes
    fig_height: 4
    fig_width: 6
---

##Overview
Part II of the assignment invovles analyzing the ToothGrowth data in the R datasets
package.  The analysis involves providing basic summary of the data, exploratory data analysis, and comparing tooth growth by supplements and dosages.  Conclusions will be drawn from the analysis.

Load knitr package
```{r load knitr, echo=FALSE}
library(knitr)
```
Set the Working Directory
````{r set working directory, echo=FALSE}
opts_knit$set(root.dir = "C:/Coursera/Statistical Inference/Course Project")
````
Load the ToothGrowth Data
````{r ToothGrowth Data}
library(datasets)
data(ToothGrowth)
````

##Basic Summary of Data
````{r Data Summary}
dim(ToothGrowth)
names(ToothGrowth)
str(ToothGrowth)
````

##Data Exploration
````{r Average length per supp}
aggregate(len ~ supp, ToothGrowth, mean)
aggregate(len ~ dose, ToothGrowth, mean)
````

````{r Tooth Growth Grouped by Supplement and Dosage}
library(ggplot2)
d <- ggplot(ToothGrowth, aes(supp, len))
d + geom_boxplot(aes(fill = factor(dose))) + 
  labs(title = "Box Plot",
       subtitle = "Tooth Growth Grouped by Supplement and Dosage ",
       x = "Supp",
       Y = "Length")
````

##Assumptions - Comparing tooth growth by supplements and dosages.
1. Null Hypothesis:  The difference in the average tooth length is zero between the supplements. 
```` {r Test Null Hypothesis}
t.test(ToothGrowth$len ~ ToothGrowth$supp, paired = FALSE)
````
- The t value of 1.9153 falls within the confidence interval, therefore
we fail to reject the null hypothesis that the difference in means is
equal to 0.

2.  Test the Null Hypothesis that the difference in the average tooth 
length is zero amongst dosages. 
- There are three comparisons: .5 to 1 mg, 1 to 2 mg, and .5 to 2 mg.  
- One is shown below:

- Compare dosages .5mg and 2mg
```` {r subsetting3}
dose.5b <- subset(ToothGrowth, ToothGrowth$dose == .5, 1)
dose2b <- subset(ToothGrowth, ToothGrowth$dose == 2, 1)
dosetest3 <- t.test(dose.5b, dose2b)
dosetest3
````
##Conclusion comparing tooth length to dosage.
For each test comparing doseages, the p-value is much smaller
compared to the alpha of 0.05.  Hence we reject the Null Hypothesis
that the difference in means amongst dosage is zero.